An Invariant Framework for Conducting Reproducible Computational Science

### 2015-03-16 evolution.tex
Modifications according to the suggestions from reviewers

Given the many challenges of reproducing HEP applications, we now describe a formal framework, which if present within large collaborations, can enable application developers to share their 
application with other researchers, and for other researchers to reproduce the shared application. In effect, the framework itself should be 
be invariant to changes in who owns the application (the developer or the reproducer), where it was built and compiled. To satisfy invariance, the framework must include mechanisms for:

@@

Documentation and specification during application development can be a onerous. The preservation framework must make literate programming tools available that 
focus less on documentation, but script, integrate and execute the dependencies so that they are resolved as part of documentation. 
Automation can extend to various tasks that are necessary for ensuring repeatability such as building software, provisioning of hardware, 
validation of software against security fixes, new features, or deprecation of old software, and even monitoring the reproducibility state of a preserved application, i.e., its source code, dependencies, environment, and platform.
Automated builds and provisioning and continuous integration service can significantly lower the barrier to run the application in a new environment.

>>>>

Given the many challenges of reproducing HEP applications, we now describe an invariant framework, which if present within large collaborations, can enable application developers to share their 
application with other researchers, and for other researchers to reproduce the shared application. 
To satisfy invariance, the framework must include mechanisms for:

@@

Documentation and specification during application development can be an onerous one. The preservation framework must make programming tools available that 
focus less on documentation, but script, integrate and execute the dependencies so that they are resolved as part of documentation. 
Automation can extend to various tasks that are necessary for ensuring repeatability such as building software, provisioning of hardware, 
validation of software against security fixes, new features, and even monitoring the reproducibility state of a preserved application, i.e., its source code, dependencies, environment, and platform.
Automated builds and provisioning and continuous integration service can significantly lower the barrier to run the application in a new environment.

### 2015-03-16 intro.tex
Modifications according to the suggestions from reviewers

% Reproducibility
Reproducibility is a cornerstone of the scientific method~\cite{borgman2012data}. 
Its importance is underscored by its ability to advance science---reproducing by verifying and validating a scientific result leads to improved understanding, thus increasing possibilities of reusing or extending the result. 
Ensuring reproducibility of a scientific result, however, often entails detailed documentation and specification of the involved scientific method. Historically, this has been achieved through text and proofs in a publication. 
As computation pervades the sciences and transforms the scientific method, mere text is considered insufficient. 
In particular, apart from textual descriptions describing the result, a reproducible result must also include several computational artifacts, such as software, data,  environment variables, and state of computation that are involved in the adopted scientific method \cite{Sole}.  By including computational artifacts, the scientific result can be obtained again at a later time.

Virtualization has emerged as a promising approach for reproducing computational scientific results. One approach is to conduct the entire computation relating to a scientific result within a virtual machine image, and then share the resulting image. This way VMIs become an authoritative, encapsulated, and executable records of computations, especially computations whose results are destined for publication and/or re-use, and the VMIs can be shared easily \cite{Lampoudi}. 
With this approach, oftentimes, the resulting image is too big in size for share-ability purposes. An alternative light-weight form of virtualization allows encapsulation of the application software, along with all its necessary dependencies into a self-contained package, but does not make it executable.  It, thus, separates and does not include the underlying operating system on which the package is executed. The encapsulation of the self-contained package is achieved by interposing application system calls, and only copying the necessary dependencies (data, libraries, code, etc) into the package, which also makes it lighter weight than a VMI. While both forms of virtualization provide mechanisms for sandboxing the computations associated with a scientific result, neither form of virtualization provides any guarantee that the included pieces of software will indeed reproduce the associated scientific result.

Since reproducibility is significantly about comprehensive documentation, virtualization approaches in their current form only make it easy to capture the computations. Preserving the computations, so that they are easy to understand, install, or alter implicit dependencies that are part of computation, especially as dependencies and software components evolve or become deprecated, is not effectively addressed. There are two approaches to address the preservation challenge. By either introducing tools that help document dependencies and provide software attribution within VMIs or packages, or alternatively by using software delivery mechanisms, such as, centralized package management, linux containers, and the, more recent, Docker framework. We examined the first approach previously in \cite{SoftProv}. In this paper we examine the second approach, that is, examine how the documentation and preservation standards that are part of software delivery mechanisms can be combined with traditional virtualization approaches for reproducibility. In particular, we consider the light-weight virtualization approaches, because, we believe that the combination of light-weight approaches with more standardized software delivery mechanisms can lead to addressing the reproducibility challenge for a wide variety of scientific researchers. 
% they may help the This  We that must also include effective  to deliver the software to a wide audience. We examine the fire
%have recently emerged  several of the reproducible challenges that arise due to use of virtualization approaches.

To conduct a thorough examination, we consider real-world complex high energy physics (HEP) applications, independently developed by two HEP groups, that must be reproduced so that the entire high energy physics community can benefit from the analysis obtained from the application. We describe challenges in reproducing the applications, and consider the extent to which reproducibility requirements can be satisfied with light-weight virtualization approaches and software delivery mechanisms. We propose an invariant framework for computational reproducibility that combines light-weight virtualization with software delivery mechanisms for efficiently capturing, invariantly preserving and practically deploying applications. In effect, the framework uses light-weight virtualization techniques for efficiently capturing a computational environment, but preserves the environment such that it is invariant to temporal software changes that may occur in near future and can be re-executed later. We measure the performance overhead of  light-weight virtualization and software delivery approaches. 
% For later. 
%The combination imposes additional overhead due to double virtualization, but saves on context switches, since every system call is not interposed.

>>>>

% Reproducibility
Reproducibility is a cornerstone of the scientific method~\cite{borgman2012data}. 
Its importance is underscored by its ability to advance science---reproducing by verifying and validating a scientific result leads to improved understanding, thus increasing possibilities of reusing or extending the result. 
Ensuring reproducibility of a scientific result, however, often entails detailed documentation and specification of the involved scientific method. Historically, this has been achieved through text and proofs in a publication. 
As computation pervades the sciences and transforms the scientific method, mere text is considered insufficient. 
In particular, apart from textual descriptions describing the result, a reproducible result must also include several computational artifacts, such as software, data,  environment variables, and state of computation that are involved in the adopted scientific method \cite{Sole}.

Virtualization has emerged as a promising approach for reproducing computational scientific results. One approach is to conduct the entire computation relating to a scientific result within a virtual machine image, and then share the resulting image. This way VMIs become an authoritative, encapsulated, and executable records of computations, especially computations whose results are destined for publication and/or re-use, and the VMIs can be shared easily \cite{Lampoudi}. 
With this approach, oftentimes, the resulting image is too big in size for share-ability purposes. An alternative light-weight form of virtualization allows encapsulation of the application software, along with all its necessary dependencies into a self-contained package, but does not make it executable.
The encapsulation of the self-contained package is achieved by interposing application system calls, and only copying the necessary dependencies (data, libraries, code, etc.) into the package, which also makes it lighter weight than a VMI. While both forms of virtualization provide mechanisms for sandboxing the computations associated with a scientific result, neither form of virtualization provides any guarantee that the included pieces of software will indeed reproduce the associated scientific result.

Since reproducibility is significantly about comprehensive documentation, virtualization approaches in their current form only make it easy to capture the computations. Preserving the computations, so that they are easy to understand, install, or alter implicit dependencies that are part of computation, especially as dependencies and software components evolve or become deprecated, is not effectively addressed. There are two approaches to address the preservation challenge. By either introducing tools that help document dependencies and provide software attribution within VMIs or packages, or alternatively by using software delivery mechanisms, such as, centralized package management, Linux containers, and the, more recent, Docker framework. We examined the first approach previously in \cite{SoftProv}. 
In this paper we examine the second approach. 
In particular, we consider the light-weight virtualization approaches, because, we believe that the combination of light-weight approaches with more standardized software delivery mechanisms can lead to addressing the reproducibility challenge for a wide variety of scientific researchers. 
A package created by those light-weight approaches encapsulates all the necessary dependencies of an application, and can be used to repeat the application through different sandbox mechanisms, including Parrot, CDE, PTU, chroot, and Docker.
% they may help the This  We that must also include effective  to deliver the software to a wide audience. We examine the fire
%have recently emerged  several of the reproducible challenges that arise due to use of virtualization approaches.

To conduct a thorough examination, we consider real-world complex high energy physics (HEP) applications, independently developed by two HEP groups, that must be reproduced so that the entire high energy physics community can benefit from the analysis obtained from the application. We describe challenges in reproducing the applications, and consider the extent to which reproducibility requirements can be satisfied with light-weight virtualization approaches and software delivery mechanisms. We propose an invariant framework for computational reproducibility that combines light-weight virtualization with software delivery mechanisms for efficiently capturing, invariantly preserving and practically deploying applications.
We measure the performance overhead of  light-weight virtualization and software delivery approaches.

We do not claim our solution is the only way to preserve applications. Generally, there are two different approaches to preserve applications: measure the mess and force cleanliness. The first approach allows end users to construct the environment as desired, and then measures the dependencies. The latter one forces users to specify the execution environment for an application in a well organized way.
Our objective here is to preserve the mess as-is.

% For later. 
%The combination imposes additional overhead due to double virtualization, but saves on context switches, since every system call is not interposed.

### 2015-03-16 measure.tex
Minor changes
Modifications according to the suggestions from reviewers

We have constructed a first approximation of the invariant framework
by using and modifying a variety of existing technologies.  Of course,
a viable long-term strategy for reproducibility must not depend on a single
technology, but on multiple technologies that interchange compatible artifacts.
To this end, we have identified several technologies and techniques
that can implement each stage of the framework.

@@

Both Parrot and PTU generate packages that consist of plain filesystem trees representing the namespace and data of the preserved application, and can be easily transformed into whatever archive format (e.g. ZIP or TGZ) is most suitable.  This is desirable for long-term preservation that may outlive various deployment technologies.
However, neither technology yet integrates the literate programming environment
or the continuous integration environment envisioned above.  Without these techniques, the preserved packages will have diminishing value over time.

@@

The latter technology is of considerable interest today

 % Docker images
\vspace{5pt}
\noindent{\bf Linux Containers and Docker Images:} Linux Containers provide multiple isolated execution instances on top of the same kernel through OS-level virtualization. Thus they can be used to persist the captured packages into images. Using such containers, Docker now allows to preserve the image in a more user-friendly way. In particular, apart from performing Linux container (LXC) based operating system (OS) level virtualization, docker tools provides portable deployment of containers across platforms, documentation of packages in a scriptable format, and versioning of container images. The image can be preserved along with dockerfiles, which are plain text files with a literate programming environment: the computational part is similar to shell scripts and can help provision similar to other provisioning tools (e.g. Chef, Puppet) or Continuous Integration (CI) platforms (e.g. Travis CI), and the text part is for human consumption and more suited for use with a version management system such as subversion or git, which can track any changes made to the Dockerfile. Thus dockerfiles can be used to preserve the namelist of Parrot packages and provenance description in PTU. 
Docker is integrated with a continuous build environment which will check and validate the version of the software being used, and use a more recent version to build application software.

>>>>

We have constructed a first approximation of the invariant framework
by using and modifying a variety of existing technologies.  Of course,
a viable long-term strategy for reproducibility must not depend on a single
technology.
To this end, we have identified several technologies and techniques
that can implement each stage of the framework.

@@

Both Parrot and PTU generate packages that consist of plain filesystem trees representing the namespace and data of the preserved application, and can be easily transformed into whatever archive format (e.g. ZIP or TGZ) is most suitable.  This is desirable for long-term preservation that may outlive various deployment technologies.
However, neither technology yet integrates the programming environment envisioned above for the documentation purpose.  Without these techniques, the preserved packages will have diminishing value over time.

@@

 % Docker images
\vspace{5pt}
\noindent{\bf Linux Containers and Docker Images:} Linux Containers provide multiple isolated execution instances on top of the same kernel through OS-level virtualization. Thus they can be used to persist the captured packages into images. 
Using such containers, Docker now allows to preserve the image in a more user-friendly way. 
Docker also provides portable deployment of containers across platforms, documentation of packages in a scriptable format, and versioning of container images. The image can be preserved along with dockerfiles, which are text files that contain all the commands to build a Docker image: the computational part is similar to shell scripts and can help provision similar to other provisioning tools (e.g. Chef, Puppet) and the text part is for human consumption and more suited for use with a version management system such as subversion or git, which can track any changes made to the Dockerfile. Thus dockerfiles can be used to preserve the namelist of Parrot packages and provenance description in PTU. 
Docker is integrated with a continuous build environment which will check and validate the version of the software being used, and use a more recent version to build application software.

### 2015-03-16 observation.tex
Modifications according to the suggestions from reviewers

\item {\bf Configuration Complexity.} To correctly reproduce an application implies that run-time configurations and consistency checks on the available software are effective captured and preserved. 
For CMS,  {\tt scram} software management tool, is used to locate
the appropriate version of software,  set environment variables such as the PATH, run any
tool-specific configuration, and do the same for all software on which it depends. A reproducible framework must capture the working of such software management tools so that they can conduct similar
checks on a new machine.

@@

\item {\bf Dependencies for Reproducible Execution.} Capturing the necessary and sufficient dependencies that are part of an application is sufficient for repeatability, but possibly for not reproducibility.
Repeatability implies if a result depends on running program X, we must be able to run exactly X again. In reproducibility the goal is rarely limited to running
\emph{precisely} what a predecessor did. Often, the objective is to
change a parameter or a data input in order to see how the result is affected. To that end, the preservation system must capture enough of the surrounding
material to permit modifications to succeed. In Athena framework, for instance, the software library needed to include basic software programs such as make, awk, sed that can aid further development.  
Further, a better understanding of how end users will consume preserved software will help to shape how
software is preserved in the first place.
\end{itemize}

>>>>

\item {\bf Configuration Complexity.} To correctly reproduce an application implies that run-time configurations and consistency checks on the available software are effectively captured and preserved. 
For CMS,  {\tt scram} software management tool, is used to locate
the appropriate version of software,  set environment variables such as the PATH, run any
tool-specific configuration, and do the same for all software on which it depends. A reproducible framework must capture the work of such software management tools so that they can conduct similar
checks on a new machine.

@@

\item {\bf Dependencies for Reproducible Execution.} Capturing the necessary and sufficient dependencies that are part of an application is sufficient for repeatability, but possibly for not reproducibility.
Repeatability implies if a result depends on running program X, we must be able to run exactly X again. In reproducibility the goal is rarely limited to running
\emph{precisely} what a predecessor did. Often, the objective is to
change a parameter or a data input in order to see how the result is affected. To that end, the preservation system must capture enough of the surrounding
material to permit modifications to succeed. 
Further, a better understanding of how end users will consume preserved software will help to shape how
software should be preserved.
\end{itemize}

### 2015-03-16 paper.tex
Challenges
Add keywords
Application revision
Final
Modifications according to the suggestions from reviewers

\usepackage{comment}
\excludecomment{TM}

@@

\keywords{}

@@

\input {intro}
\input {tauroast}
\input {observation}
\input {evolution}
\input {measure}
\input {evaluation}
\input {problems}
\input {related_work}
%\input {postscript}

@@

\bibliographystyle{plain}
\bibliography{cclpapers,this}

>>>>

\usepackage{comment}
\usepackage{wrapfig}
\excludecomment{TM}
%\setlength{\textwidth}{6.25in}

@@

\keywords{Preservation framework, reproducible research, virtualization, container}

@@

\input {intro}
\input {tauroast}
\input {observation}
\input {evolution}
\input {measure}
\input {evaluation}
%\input {problems}
\input {related_work}
%\input {postscript}
\section{Conclusions and Future work}
In this paper, we propose an invariant framework for conducting reproducible computational science - using light-weight virtualization approaches to preserve applications in the format of self-contained packages and using standardized software delivery mechanisms to deliver and distribute preserved packages.
We use two complex high energy physics applications to illustrate how the framework can help the original authors preserve and distribute the applications, and others reproduce the applications.

This paper focuses on how to measure the mess and track the used dependencies to preserve an application.
In the following work, we plan to explore how to preserve an application in an organized style - specifying the execution environment clearly.
How to preserve and improve the availability of remote network resources is another important problem to be explored.

For more details about how to preserve, distribute and repeat an application:

\url{http://ccl.cse.nd.edu/research/data/hep-case-study/case\_study\_redhat6\_5.html}

@@

\bibliographystyle{plain}
\bibliography{cclpapers,this,sole}
